Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-08-20 papers/paper/ms.tex
update ms

We were interested in not only the overall age distribution of the
Kepler field, but also in the reliability of individual ages for young
stars known to host planets.  We attempted to assess this reliability using
quality bitmasks (Table~\ref{tab:planets}), a comparison against \deleted{open }
cluster ages (Section~\ref{sec:litagecomparison}), and a comparison
between rotation and lithium ages.  The \deleted{open }cluster comparison
suggested that our ages were generally accurate at $\ll$0.5\,dex
across 0.04--2.5\,Gyr, similar to their quoted precision.
Nonetheless, the upper-left panel of Figure~\ref{fig:agescalecompone}
does show outliers, typically $\lesssim$10\% of the population
in any cluster.  These stars could be either field star contaminants,
or else anomalous stars whose physical rotation histories were altered
by processes not captured by our statistical uncertainties.
Similarly, although we found $t_{\rm gyro}$ and $t_{\rm Li}$ to agree
for \ltonegyrhighqconfirmedtwosided\ planets orbiting ``high-quality''
stars, we did find one system, Kepler-786, with radically different
lithium (\kepseveneightsix\,Myr) and rotational
(\kepseveneightsixgyro\,Myr) ages.  Planetary mergers
are one process that could produce such a signal, but
testing this would require a method for measuring differential abundances
across a large number of elements.

>>>>

We were interested in not only the overall age distribution of the
Kepler field, but also in the reliability of individual ages for young
stars known to host planets.  We attempted to assess this reliability using
quality bitmasks (Table~\ref{tab:planets}), a comparison against \deleted{open }
cluster ages (Section~\ref{sec:litagecomparison}), and a comparison
between rotation and lithium ages.  The \deleted{open }cluster comparison
suggested that our ages were generally accurate at $\ll$0.5\,dex
across 0.04--2.5\,Gyr, similar to their quoted precision.
Nonetheless, the upper-left panel of Figure~\ref{fig:agescalecompone}
does show outliers, typically $\lesssim$10\% of the population
in any cluster.  These stars could be either field star contaminants,
or else anomalous stars whose physical rotation histories were altered
by processes not captured by our statistical uncertainties.
Similarly, although we found $t_{\rm gyro}$ and $t_{\rm Li}$ to agree
for \ltonegyrhighqconfirmedtwosided\ planets orbiting ``high-quality''
stars, we did find one system, Kepler-786, with radically different
lithium (\kepseveneightsix\,Myr) and rotational
($\approx$4.4\,Gyr) ages.  Planetary mergers
are one process that could produce such a signal, but
testing this would require a method for measuring differential abundances
across a large number of elements.
